---
title: "Imputation of Missing Pollution Data"
author: "Phillip Sangwoo Lee"
output:
  html_document: default
---

```{r load-pkg, message = FALSE}
library(modelsummary)
library(mice)
library(brms)
library(magrittr)
library(posterior)
library(stargazer)
library(dplyr)
library(tidyr)
library(haven)
library(bayesplot)
library(tinytex)
```

Import the data set using the following code (and make changes as needed). 

```{r dat}
pm <- read_dta("/Users/philee/Dropbox/USC/2022 Winter/Bayesian/Final Paper/bayes_final(daily).dta")
pm$Station_ID <- as.factor(pm$Station_ID)
pm$month <- as.factor(pm$month)
pm$year <- as.factor(pm$year)

pm <- pm[c("myday", "Station_ID", "month", "year", "davsr", "davrh", "daily_avg_pm10", "dtmp8m", "ftot", "fmax")]
```

Summary Statistics of Observed vs Missing 

```{r PM10_missing, indent = "    "}
datasummary_balance(
    ~ missing_mp,
    data = cbind(pm[ , -c(1:4)],
                 missing_mp = factor(is.na(pm$daily_avg_pm10),
                                     labels = c("observed", "missing"))))
```

Summary Statistics of Observed vs Missing by Station

```{r}
pm %>% 
  mutate(missings = ifelse(is.na(daily_avg_pm10), "missing", "observed")) %>% 
  select(myday, Station_ID, missings, dtmp8m, davsr, davrh, ftot, fmax) %>% 
  group_by(Station_ID, missings) %>% 
  mutate(id = 1:n()) %>% 
  ungroup() %>% 
  gather(temp, val, dtmp8m, davsr, davrh, ftot, fmax) %>% 
  unite(temp1, temp, Station_ID, missings, sep = "_") %>% 
  spread(temp1, val) %>% 
  select(-id) %>% 
  as.data.frame() %>% 
  stargazer(type = "text")
```

From the missing data pattern below, how many measurements have missing PM10 and PM2.5?

```{r missing-pattern, indent = "    ", results = "hide", fig.asp = 1.618}
mice::md.pattern(pm, rotate = TRUE)
sum(is.na(pm$daily_avg_pm10)) + sum(is.na(pm$daily_avg_pm25)) - sum(is.na(pm$daily_avg_pm10)&is.na(pm$daily_avg_pm25))
sum(is.na(pm$daily_avg_pm10)|is.na(pm$daily_avg_pm25))
```

Single Imputation Model

```{r m1, indent = "    ", results = "hide", message = FALSE}
m1 <- brm(daily_avg_pm10 | mi() ~ davsr + davrh + dtmp8m + ftot + fmax + month + year + (1|Station_ID), data = pm,
          prior = c(
              prior(normal(18,10), class = "Intercept"),
              prior(normal(169,46), class = "b", coef = davsr),
              prior(normal(65,6), class = "b", coef = davrh),
              prior(normal(13,15), class = "b", coef = dtmp8m),
              prior(normal(538,857), class = "b", coef = ftot),
              prior(normal(104,107), class = "b", coef = fmax)),
          chains = 4, cores = 8, iter = 4000, seed = 1)
```

Single Imputation Model Estimates

```{r m1 model estimates}
m1
```

5 Simulated/Imputed Datasets

```{r imp, indent = "    "}
imp <- mice(pm, maxit = 0)
pred <- imp$predictorMatrix
# Exclude id
pred[ , c("myday", "Station_ID", "daily_avg_pm10", "daily_avg_pm25")] <- 0
# imputation 
imp <- mice(pm, m = 5, predictorMatrix = pred, 
            maxit = 50, 
            print = FALSE,
            seed = 111)
```

Check for mixture of simulated datasets

```{r plot-imp, indent = "    ", out.width = "100%", fig.width = 8.57}
plot(imp)
```

Multiple Imputation Model Estimates with 50 Imputed Datasets

```{r m2, indent = "    ", results = "hide", message = FALSE}
m2 <- brm_multiple(
    daily_avg_pm10 ~ davsr + davrh + dtmp8m + ftot + fmax + month + year + (1|Station_ID), data = imp,
          prior = c(
              prior(normal(18,10), class = "Intercept"),
              prior(normal(169,46), class = "b", coef = davsr),
              prior(normal(65,6), class = "b", coef = davrh),
              prior(normal(13,15), class = "b", coef = dtmp8m),
              prior(normal(538,857), class = "b", coef = ftot),
              prior(normal(104,107), class = "b", coef = fmax)),
    chains = 4, cores = 8, iter = 4000, seed = 1) 
```

```{r m2-plot, indent = "    "}
mcmc_plot(m2, type = "dens", variable = "b_Intercept")
```


```{r}
m2 
```

```{r}
as_draws_df(m2) %>%
    subset_draws(chain = 1:4) %>%
    summarize_draws()
```


```{r plot m2}
plot(m2)
```


50 Simulated/Imputed Datasets

```{r imp, indent = "    "}
imp2 <- mice(pm, maxit = 0)
pred <- imp2$predictorMatrix
# Exclude id
pred[ , c("myday", "Station_ID", "daily_avg_pm10", "daily_avg_pm25")] <- 0
# imputation 
imp2 <- mice(pm, m = 50, predictorMatrix = pred, 
            maxit = 50, 
            print = FALSE,
            seed = 111)
```

Check for mixture of simulated datasets

```{r plot-imp, indent = "    ", out.width = "100%", fig.width = 8.57}
plot(imp2)
```


Multiple Imputation Model Estimates with 50 Imputed Datasets

```{r m2, indent = "    ", results = "hide", message = FALSE}
m3 <- brm_multiple(
    daily_avg_pm10 ~ davsr + davrh + dtmp8m + ftot + fmax + month + year + (1|Station_ID), data = imp2,
          prior = c(
              prior(normal(18,10), class = "Intercept"),
              prior(normal(169,46), class = "b", coef = davsr),
              prior(normal(65,6), class = "b", coef = davrh),
              prior(normal(13,15), class = "b", coef = dtmp8m),
              prior(normal(538,857), class = "b", coef = ftot),
              prior(normal(104,107), class = "b", coef = fmax)),
    chains = 4, cores = 8, iter = 4000, seed = 1) 
```

```{r m2-plot, indent = "    "}
mcmc_plot(m3, type = "dens", variable = "b_Intercept")
```


```{r}
m3 
```

```{r}
as_draws_df(m3) %>%
    subset_draws(chain = 1:4) %>%
    summarize_draws()
```


```{r plot m3}
plot(m3)
```