---
title: "Homework 9"
author: "Naomi Seula Lee"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    math_method: webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-pkg, message = FALSE}
library(haven)   # for opening dta file
library(ggplot2)  # for plots
library(magrittr)  # for `%>%` operator
library(here)
library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(rstan)
rstan_options(auto_write = TRUE)  # save compiled STAN object
options(mc.cores = 2)  # use two cores
library(posterior)
library(bayesplot)
theme_set(theme_classic() +
    theme(panel.grid.major.y = element_line(color = "grey92")))
```

# Research Question

> Do we observe higher PM readings for months in which forest fires were frequently observed?

For our analysis project, we aim to estimate the missing values for Particulate Matter(PM) records in Bogota, Columbia. In Bogota, Columbia, forest fire is prevalent in the neighboring regions, and such forest fire frequently happens in February and March. Forest fire is expected to increase the PM concentrate measure, and thus, we first want to check whether we observe higher PM readings for the months with frequent forest fire.

# Variables

- `pm10`: Particulate matter 10 (micrograms/$m^3$)
- `pm25`: Particulate matter 2.5 (micrograms/$m^3$)
- `myday`: Day of the year in format ddmmyyyy
- `month`: Month of the year
- `rain`: Rainfall (mm)
- `Station_ID`: Pollutant tracking station ID
- `tmp`: Temperature at 20m from ground level (degrees in Celsius)
- `quadrant`: Quadrant in which the station is located (1 = NE, 2 = SE, 3 = SW, 4 = NW)
- `tot`: Total number of forest fires within 1,500km X 1,500km from the center of Bogota, Columbia
- `qtot`: `tot` by quadrant
- `max`: Daily max forest Fire Radiative Power (FRP)
- `qmax`: `tot` by quadrant
- `fire`: A dummy variable indicating months that forest fires were frequently observed

## Import Data

```{r data import}
data <- read_dta("/Users/seulalee/Desktop/Spring 2022/PSYC 573/Project/bayes_final1.dta")
```

## Variable Summary

```{r summ-var}
datasummary(daily_avg_pm10  * 
                (N + Mean + SD + Min + Max + Histogram) ~
                factor(fire, labels = c("Non_Fire", "Fire")),
            data = data)

datasummary(daily_avg_pm25  * 
                (N + Mean + SD + Min + Max + Histogram) ~
                factor(fire, labels = c("Non_Fire", "Fire")),
            data = data)
```

From the variable summary, we can see that months with frequent forest fire (fire=1) has higher records for both PM10 and PM2.5 measure. In the following part, we will use PM10 values to see whether there actually exist meaningful differences between months with and without frequent forest fire.

# Model

Let $Y$ = PM10, $F$ = Forest Fire

Model: $$
  \begin{aligned}
    Y_{i, F = 0} & \sim N(\mu_1, \sigma_1) \\
    Y_{i, F = 1} & \sim N(\mu_2, \sigma_2)
  \end{aligned}
$$

Prior: $$
  \begin{aligned}
    \mu_1 & \sim N(36, 20) \\
    \mu_2 & \sim N(45, 23) \\
    \sigma_1 & \sim N^+(0, 2) \\
    \sigma_2 & \sim N^+(0, 2)
  \end{aligned}
$$

## Running Stan

We used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups).

```{r m1, results = "hide"}
# 1. form the data list for Stan
stan_dat <- with(data,
    list(N1 = sum(fire == 0),
         N2 = sum(fire == 1),
         y1 = daily_avg_pm10[which(fire == 0)],
         y2 = daily_avg_pm10[which(fire == 1)])
)
# 2. Run Stan
m1 <- stan(
    file = here("Project", "normal_2group.stan"),
    data = stan_dat,
    seed = 2834,  # for reproducibility
    iter = 4000
)
```

# Results

```{r}
print(m1, pars = c("mu1", "mu2", "sigma1", "sigma2"))
```

As shown in the graph below, the chains mixed well.

```{r rank-hist-m1}
mcmc_rank_hist(m1, pars = c("mu1", "mu2", "sigma1", "sigma2"))
```

```{r summ-m1}
summ_m1 <- as_draws_df(m1) %>%
    subset_draws(variable = c("mu1", "mu2", "sigma1", "sigma2")) %>%
    mutate_variables(`mu2 - mu1` = mu2 - mu1) %>%
    summarise_draws()
knitr::kable(summ_m1, digits = 2)
```

The following table shows the posterior distributions of $\mu_1$, $\mu_2$, $\sigma_1$, $\sigma_2$, and $\mu_2 - \mu_1$. As the last row `mu2-mu1` value suggest, the months with frequent forest fire has higher PM readings than the months without frequent forest fire with the difference of `r round(summ_m1$mean[5], 2)`

```{r mcmc_dens}
mcmc_dens(m1, 
          pars = c("mu1","sigma1","mu2","sigma2"))
```

The analysis showed that on average, months with frequent forest fire showed higher PM concentrate values than the months that did not observe frequent forest fire, with a posterior mean of `r round(summ_m1$mean[5], 2)` and a 90% CI of [`r round(summ_m1$q5[5], 2)`, `r round(summ_m1$q95[5], 2)`].
