---
title: "homework 9"
author: "Seongmoon Cho"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-pkg, message = FALSE}
library(haven)   # for opening dta file
library(ggplot2)  # for plots
library(magrittr)  # for `%>%` operator
library(here)
library(readxl)  # for reading excel files
library(modelsummary)  # for summarizing data
library(rstan)
rstan_options(auto_write = TRUE)  # save compiled STAN object
options(mc.cores = 2)  # use two cores
library(posterior)
library(bayesplot)
theme_set(theme_classic() +
    theme(panel.grid.major.y = element_line(color = "grey92")))
```

# Research Question

> Do we observe higher PM readings for months in which forest fires were frequently observed?


# Variables

- `pm10`: Particulate matter 10 (micrograms/$m^3$)
- `pm25`: Particulate matter 2.5 (micrograms/$m^3$)
- `myday`: Day of the year in format ddmmyyyy
- `month`: Month of the year
- `rain`: Rainfall (mm)
- `Station_ID`: Pollutant tracking station ID
- `tmp`: Temperature at 20m from ground level (degrees in Celsius)
- `quadrant`: Quadrant in which the station is located (1 = NE, 2 = SE, 3 = SW, 4 = NW)
- `tot`: Total number of forest fires within 1,500km X 1,500km from the center of Bogota, Columbia
- `qtot`: `tot` by quadrant
- `max`: Daily max forest Fire Radiative Power (FRP)
- `qmax`: `tot` by quadrant
- `fire`: A dummy variable indicating months that forest fires were frequently observed

## Import Data

```{r data import}
data <- read_dta(here("bayes_final1.dta"))
```

## Variable Summary

```{r summ-var}
datasummary(daily_avg_pm10  * 
                (N + Mean + SD + Min + Max + Histogram) ~
                factor(fire, labels = c("Non_Fire", "Fire")),
            data = data)
```

# Model

Let $Y$ = PM10, $G$ = Forest Fire

Model:
$$
  \begin{aligned}
    Y_{i, G = 0} & \sim N(\mu_1, \sigma_1) \\
    Y_{i, G = 1} & \sim N(\mu_2, \sigma_2)
  \end{aligned}
$$

Prior:
$$
  \begin{aligned}
    \mu_1 & \sim N(36, 20) \\
    \mu_2 & \sim N(45, 23) \\
    \sigma_1 & \sim N^+(0, 2) \\
    \sigma_2 & \sim N^+(0, 2)
  \end{aligned}
$$

## Running Stan

We used 4 chains, each with 4,000 iterations (first 2,000 as warm-ups). 

```{r m1, results = "hide"}
# 1. form the data list for Stan
stan_dat <- with(data,
    list(N1 = sum(fire == 0),
         N2 = sum(fire == 1),
         y1 = daily_avg_pm10[which(fire == 0)],
         y2 = daily_avg_pm10[which(fire == 1)])
)
# 2. Run Stan
m1 <- stan(
    file = here("normal_2group.stan"),
    data = stan_dat,
    seed = 20220419,  # for reproducibility
    iter = 4000
)
```

# Results
```{r}
print(m1, pars = c("mu1", "mu2", "sigma1", "sigma2"))
```


As shown in the graph below, the chains mixed well.

```{r rank-hist-m1}
mcmc_rank_hist(m1, pars = c("mu1", "mu2", "sigma1", "sigma2"))
```

The following table shows the posterior distributions of $\mu_1$, $\mu_2$, $\sigma_1$, $\sigma_2$, and $\mu_2 - \mu_1$.

```{r summ-m1}
summ_m1 <- as_draws_df(m1) %>%
    subset_draws(variable = c("mu1", "mu2", "sigma1", "sigma2")) %>%
    mutate_variables(`mu2 - mu1` = mu2 - mu1) %>%
    summarise_draws()
knitr::kable(summ_m1, digits = 2)
```

```{r mcmc_dens}
mcmc_dens(m1, 
          pars = c("mu1","sigma1","mu2","sigma2"))
```

The analysis showed that on average, months with frequent forest fire showed higher PM concentrate than months that did not observe frequent forest fire, with a posterior mean of `r round(summ_m1$mean[5], 2)` and a 90% CI of [`r round(summ_m1$q5[5], 2)`, `r round(summ_m1$q95[5], 2)`].